---
title: "Exercise1"
author: "Dario Pedolin"
date: "26 2 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require("knitr")){
  install.packages("knitr")
  require("knitr")
}
if(!require("tidyverse")){
  install.packages("tidyverse")
  require("tidyverse")
}
if(!require("openxlsx")){
  install.packages("openxlsx")
  require("openxlsx")
}



source("../../util/functions.R")

```




## 1.1 Goal and system defintion

### System boundaries
  
  * Spatial reference: Europe
  * Time reference: 2019 - 2050
  * Substantial reference: E-Bikes

### Guiding questions
	* Which impact factors and interrelations among them determine…
	  * The penetration of different types of e-bikes and battery technologies on the EU market?
	  * How the adoption of e-bikes changes the mobility behavior and what modes of transport are being replaced (substitution) to which extent?
	  * in Europe in the year 2050.
	* What scenarios can result from different constellations of the identified impact factors?
	
	By the way:
	* What other consequences could a large scale adoption of e-bikes have?
	* What factors could lead to a rebound effect?

### Knowledge base
	-
	-
### Actors to be involved
	-
### Target group
	-


## 2.1 Impact factors
### preliminary identification:
```{r prelimIF, echo=FALSE}
# load csv and write its content to excel
if(FALSE & file.exists("../../data/prelimImpactFactors.csv")){
  prelimImpactFactors <- read.csv("../../data/prelimImpactFactors.csv", stringsAsFactors = FALSE)
  prelimImpactFactors$Indicator <- NA
  prelimImpactFactors$Current_Stat <- NA
  
  dataXlsxName <- "../../data/exerciseData.xlsx"
  if(!file.exists(dataXlsxName)){
    write.excel(data = prelimImpactFactors, file = "../../data/exerciseData.xlsx", sheetName = "preliminaryImpactFactors", rNames = FALSE)
  } 
  if(file.exists(dataXlsxName) & "preliminaryImpactFactors" %in% excel.sheets(dataXlsxName)){
    write.excel(data = prelimImpactFactors, file = "../../data/exerciseData.xlsx", sheetName = "preliminaryImpactFactors", rNames = FALSE, append = TRUE, overwrite = TRUE)
  }
  if(file.exists(dataXlsxName) & !("preliminaryImpactFactors" %in% excel.sheets(dataXlsxName))){
    write.excel(data = prelimImpactFactors, file = "../../data/exerciseData.xlsx", sheetName = "preliminaryImpactFactors", rNames = FALSE, append = TRUE)
  }
}
# and use the excel
prelimImpactFactors <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "preliminaryImpactFactors")

prelimImpactFactorsList <- as.character(prelimImpactFactors$ImpactFactor)
cols <- 3
nrowPerCol <- ceiling(length(prelimImpactFactorsList)/cols)
prelimImpFac <- c(prelimImpactFactorsList, rep("", each = cols * nrowPerCol - length(prelimImpactFactorsList)))
prelimImpFacTab <- cbind(prelimImpFac[1:nrowPerCol], prelimImpFac[(nrowPerCol+1):(2*nrowPerCol)], prelimImpFac[(2*nrowPerCol+1):(3*nrowPerCol)])
kable(prelimImpFacTab, caption = "Names of preliminary impact factors")


```
		

	Impact factors structuring, clustering, relevance assessment
	-
	-
	
	Impact factors selection
```{r finalIF, echo=FALSE}
finalmpactFactors <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "preliminaryImpactFactors")
finalmpactFactorsList <- as.character(finalmpactFactors$ImpactFactor)
cols <- 3
nrowPerCol <- ceiling(length(finalmpactFactorsList)/cols)
impFac <- c(finalmpactFactorsList, rep(NA, each = cols * nrowPerCol - length(finalmpactFactorsList)))
impFacTab <- cbind(impFac[1:nrowPerCol], impFac[(nrowPerCol+1):(2*nrowPerCol)], impFac[(2*nrowPerCol+1):(3*nrowPerCol)])
# kable(impFacTab, caption = "Names of seleceted impact factors")
kable(finalmpactFactors, caption="Description of seleceted impact factors")
```
	-
	-
	
## 2.2 Impact assessment
	m x m matrix of direct impacts
	
```{r updateImpactMatrix, echo=FALSE, warning=FALSE}

mkeIFM <- function(IFList){
  iMatrix <- matrix(0, nrow = length(IFList), ncol=length(IFList))
  rownames(iMatrix) <- IFList
  colnames(iMatrix) <- IFList
  impactMatrix <- as.data.frame(iMatrix)
  for(i in 1:nrow(impactMatrix)){
    impactMatrix[i,i] <- NA
  }
}


finalmpactFactorsList <- as.character(finalmpactFactors$ImpactFactor)

dataXlsxName <- "../../data/exerciseData.xlsx"
if(file.exists(dataXlsxName) & !("impactMatrix" %in% excel.sheets(dataXlsxName))){
  impactMatrix <- mkeIFM(finalmpactFactorsList)
  write.excel(data = impactMatrix, file = "../../data/exerciseData.xlsx", sheetName = "impactMatrix", append = TRUE, rNames = TRUE)
}

# write updatet matrix if there are new impactFactors
if(file.exists("../../data/exerciseData.xlsx") & ("impactMatrix" %in% excel.sheets(dataXlsxName))){
  oldImpactMatrix <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "impactMatrix")
  rownames(oldImpactMatrix) <- oldImpactMatrix[,1]
  oldImpactMatrix <- oldImpactMatrix[,-1]
  existingIFIndex <- which(rownames(oldImpactMatrix) %in% finalmpactFactorsList)
  newIFNames <- finalmpactFactorsList[!(finalmpactFactorsList %in% rownames(oldImpactMatrix))]
  newImpactMatrix <- oldImpactMatrix[existingIFIndex, existingIFIndex]
  if(length(newImpactMatrix) == 0){
    newImpactMatrix == mkeIFM(finalmpactFactorsList)
  } else {
    for(newIF in newIFNames) {
      newData <- 0
      newImpactMatrix <- cbind(newImpactMatrix, newData)
      colnames(newImpactMatrix)[ncol(newImpactMatrix)] <- newIF
      newImpactMatrix <- rbind(newImpactMatrix, newData)
      rownames(newImpactMatrix)[nrow(newImpactMatrix)] <- newIF
      newImpactMatrix[nrow(newImpactMatrix), ncol(newImpactMatrix)] <- NA
    }
  }
  write.excel(data = newImpactMatrix, file = "../../data/exerciseData.xlsx", sheetName = "impactMatrix", append = TRUE, overwrite = TRUE, rNames = TRUE)

}
```

```{r impactMatrix, echo=FALSE, warning=FALSE}
# always load from excel
impactMatrix <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "impactMatrix")
names(impactMatrix)[1] <- "ImpactFactor"
impactMatrixPlot <- impactMatrix %>% 
  gather(Dependend, value, -ImpactFactor) %>% 
  mutate(value = factor(value)) %>% 
  mutate(value = factor(value),
         ImpactFactor = factor(ImpactFactor),
         ImpactFactor = factor(ImpactFactor, levels = rev(levels(ImpactFactor)))) %>%
  filter(!is.na(value)) %>% 
  ggplot(aes(Dependend, ImpactFactor, fill=value)) +
  geom_tile() +
  scale_fill_manual(values = c("0" = mc(5), "1" = mc(4),"2" = mc(1)), na.value=mc(7))+
  geom_text(aes(label = value)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

impactMatrixPlot
  
```
	
## 2.3 Impact analysis

```{r activityPassivityScores, echo=FALSE}
impactM <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "impactMatrix")
rownames(impactM) <- impactM[,1]
impactM <- impactM[,-1]
impactM$ActivityScore <- rowSums(impactM, na.rm = TRUE)
PassivityScore <- colSums(impactM, na.rm = TRUE)
impactM <- rbind.data.frame(impactM, PassivityScore= PassivityScore)
names(impactM) <- gsub(".", " ", names(impactM), fixed=TRUE)

kable(impactM)

```

### System grid
```{r plotSystemGrid, echo=FALSE}
impactM <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "impactMatrix")
rownames(impactM) <- impactM[,1]
impactM <- impactM[,-1]
impactM$ActivityScore <- rowSums(impactM, na.rm = TRUE)
PassivityScore <- colSums(impactM, na.rm = TRUE)
impactM <- rbind.data.frame(impactM, PassivityScore= PassivityScore)

sysGridData <- impactM
sysGridData$PassivityScore <- as.numeric(t(sysGridData[nrow(sysGridData),]))
sysGridData$ImpactFactor <- rownames(sysGridData)
sysGridData <- sysGridData %>%
  select(ImpactFactor, ActivityScore, PassivityScore) %>%
  filter(!(ImpactFactor %in% c("PassivityScore")))
  
library(ggrepel)

sysGridData %>% 
  ggplot(aes(ActivityScore, PassivityScore, label=ImpactFactor)) +
  geom_point() +
  geom_hline(aes(yintercept = mean(PassivityScore))) +
  geom_vline(aes(xintercept = mean(ActivityScore))) +
  geom_text_repel()


```

### Feedback loops
###System structure
	
	
## 3.1 Future level definition
## 3.2 Consistency assessment
## 3.3 Scenario construction
## 4.1 Scenario selection
## 4.2 Scenario description and interpretation
