---
title: "Exercise1"
date: "26 2 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require("knitr")){
  install.packages("knitr")
  require("knitr")
}
if(!require("tidyverse")){
  install.packages("tidyverse")
  require("tidyverse")
}
if(!require("openxlsx")){
  install.packages("openxlsx")
  require("openxlsx")
}
if(!require("htmltab")){
  install.packages("htmltab")
  require("htmltab")
}
if(!require("tabulizer")){
  install.packages("tabulizer")
  require("tabulizer")
}
if(!require("stringr")){
  install.packages("stringr")
  require("stringr")
}
if(!require("ggrepel")){
  install.packages("ggrepel")
  require("ggrepel")
}

source("../../util/functions.R")

```




## 1.1 Goal and system defintion

### System boundaries
  
  * Spatial reference: Europe
  * Time reference: 2019 - 2050
  * Substantial reference: E-Bikes

### Guiding questions
	* Which impact factors and interrelations among them determine…
	  * The penetration of different types of e-bikes and battery technologies on the EU market?
	  * How the adoption of e-bikes changes the mobility behavior?
	  * What modes of transport are being replaced (substitution) to which extent?
	  * in Europe in the year 2050.
	* What scenarios can result from different constellations of the identified impact factors?
	
	By the way:
	* What other consequences could a large scale adoption of e-bikes have?
	* What factors could lead to a rebound effect?
	


### Knowledge base
```{r loadSomeData, echo=FALSE, warning=FALSE}
if(FALSE){
  
  # Annual statistics velosuisse:
  # http://www.velosuisse.ch/de/statistik_aktuell.html
  velosuisseNeuverkaeufe <- extract_tables("http://www.velosuisse.ch/files/Veloverkauf-Statistik_2017.pdf")
  velosuisseGesamtmarkt <- as.data.frame(velosuisseNeuverkaeufe[[1]][,c(2,3,9)])
  names(velosuisseGesamtmarkt) <- c("Segment", "Total", "Diff")
  write.excel(data = velosuisseGesamtmarkt, file = "../../data/knowledgebase.xlsx", sheetName = "velosuisseGesamtmarkt", append = TRUE, overwrite = TRUE)
  
  
  # Verbreitung und Auswirkungen von E-Bikes in der Schweiz:
  # http://www.news.admin.ch/NSBSubscriber/message/attachments/36764.pdf
  # Abbildung 2-15 page 49
  distributionEffectsEBikesCH <- extract_tables("http://www.news.admin.ch/NSBSubscriber/message/attachments/36764.pdf")
  factorsPrimEnergyGHGPotential <- as.data.frame(distributionEffectsEBikesCH[[24]])
  factorsPrimEnergyGHGPotential[,2:3] <- apply(factorsPrimEnergyGHGPotential[,2:3], FUN=as.numeric, MARGIN = 2)
  names(factorsPrimEnergyGHGPotential) <- c("mode", "primary energy [MJ equ. per pkm]", "GHG potential [CO2 eq. per pkm]")
  write.excel(data = factorsPrimEnergyGHGPotential[c(-1,-2),], file = "../../data/knowledgebase.xlsx", sheetName = "factorsPrimEnergyGHGPotential", append = TRUE, overwrite = TRUE)
  
  # Abbildung 3-45, 3-46 page 105, 106
  effectEBikeVSregularBike <- data.frame(group = c("viel seltener", "seltener", "etwa gleich oft", "oefter", "viel oefter", "nie genutzt"),
                                         FahrradTouren = c(0.44, .11, .06,.02,0.03,.34),
                                         AutoArbeitsweg = c(0.32,.36, .12,.01, 0, .19),
                                         MotorradArbeitsweg = c(.34,.25, .19, 0, 0, .22),
                                         KleinmotorradArbeitsweg = c(.25,.19,.13,0,0,.44),
                                         MofaArbeitsweg = c(.2,.1,.1,0, 0,.6),
                                         Bike = c(.37,.25,.13,.01,0.01,.23),
                                         Walk = c(.13,.15,.25,.02,0,.44),
                                         OeV = c(.22,.26,.23,.01,.02,.25))
  write.excel(data = effectEBikeVSregularBike, file = "../../data/knowledgebase.xlsx", sheetName = "effectEBikeVSregularBike", append = TRUE, overwrite = TRUE)
  
  
  # Webpage European Cyclist Federation (ECF): https://ecf.com/
  ecfCapitalCities <- htmltab("https://ecf.com/resources/cycling-facts-and-figures/capital-cities", which = "/html/body/div[2]/div[3]/div/div[2]/div/section/div/div[2]/div/div[1]/div[1]")
  names(ecfCapitalCities) <- gsub("\\.*$", "", make.names(names(ecfCapitalCities)))
  write.excel(data = ecfCapitalCities, file = "../../data/knowledgebase.xlsx", sheetName = "ecfCapitalCities", append = TRUE, overwrite = TRUE)
  
  
  # blocked by admin, pdf not well foramtted
  # European e-bike market outlook:
  # http://www.biketaiwan.com/resource/article/6/157/article-08.pdf
  # euEBikeOutlook <- extract_tables("http://www.biketaiwan.com/resource/article/6/157/article-08.pdf")
  # euBike <- as.data.frame()
  
  
  # broken link
  # European bicycle market (2016 edition): 
  # http://www.conebi.eu/wpcontent/uploads/2016/09/European-Bicycle-Industry-and-Market- Profile-2016-with-2015-data-.pdf
  # euBikeMarket2016 <- extract_tables("http://www.conebi.eu/wpcontent/uploads/2016/09/European-Bicycle-Industry-and-Market- Profile-2016-with-2015-data-.pdf")

 
  # European bicycle market analysis 2015:
  # https://ecf.com/sites/ecf.com/files/CONEBI%20market%20report%20analysis%202016_1.pdf
  # Table 1 page 17
  ecfMarkteAnalysis <- extract_tables("https://ecf.com/sites/ecf.com/files/CONEBI%20market%20report%20analysis%202016_1.pdf")
  ecfMarketEU <- as.data.frame(ecfMarkteAnalysis[[2]][c(-1,-2,-3),c(1,2,3,5,8)], stringsAsFactors = FALSE)
  names(ecfMarketEU) <- paste0(ecfMarkteAnalysis[[2]][1,],ecfMarkteAnalysis[[2]][2,],ecfMarkteAnalysis[[2]][3,])[c(-5,-6,-8,-9)]
  ecfMarketEU[,2:5] <- apply(ecfMarketEU[,2:5], FUN = gsub, MARGIN = 2, pattern = " ", replacement = "")
  ecfMarketEU[,2:5] <- apply(ecfMarketEU[,2:5], FUN = as.numeric, MARGIN = 2)
  write.excel(data = ecfMarketEU, file = "../../data/knowledgebase.xlsx", sheetName = "ecfMarketEU", append = TRUE, overwrite = TRUE)
  
  
  # broken link
  # globalEBikeMarket: <- extract_tables("http://www.insg.org/docs/INSG_Insight_23_Global_Ebike_Market.pdf")
  
  
  # Electrification of road transport – an analysis of the economic performance of electric twowheelers:
  # https://dspace.library.uu.nl/bitstream/handle/1874/275936/Thesis%20P.W.K.%20Dekker%2012%20May%202013.pdf?sequence=1
  # Table 10 page 50
  economicPerformance <- extract_tables("https://dspace.library.uu.nl/bitstream/handle/1874/275936/Thesis%20P.W.K.%20Dekker%2012%20May%202013.pdf?sequence=1")
  pricesConventionalBike <- as.data.frame(apply(economicPerformance[[20]][c(-1,-2),], FUN = as.numeric, MARGIN = 2))[,c(1,3,5,7,9)]
  names(pricesConventionalBike) <- c("Year",paste0(economicPerformance[[20]][1,],economicPerformance[[20]][2,])[c(2,4,6,8)])
  write.excel(data = pricesConventionalBike, file = "../../data/knowledgebase.xlsx", sheetName = "pricesConventionalBike", append = TRUE, overwrite = TRUE)
  
  #  Table 12, 13 page 50 -52
  priceBatteryPower <- as.data.frame(economicPerformance[[21]][c(-1,-2),],stringsAsFactors = FALSE)
  names(priceBatteryPower) <- paste0(economicPerformance[[21]][1,],economicPerformance[[21]][2,])
  priceBatteryPower$Year <- 2011
  priceBatteryPower2012 <- as.data.frame(cbind(Year = 2012, economicPerformance[[22]]),stringsAsFactors = FALSE)
  names(priceBatteryPower2012) <- names(priceBatteryPower) 
  priceBatteryPower20112012 <- rbind(priceBatteryPower2012, priceBatteryPower)
  priceBatteryPower20112012$Year <- as.numeric(priceBatteryPower20112012$Year)
  priceBatteryPower20112012$`Price€` <- as.numeric(priceBatteryPower20112012$`Price€`)
  priceBatteryPower20112012$`Battery powerkWh` <- as.numeric(priceBatteryPower20112012$`Battery powerkWh`)
  write.excel(data = priceBatteryPower20112012, file = "../../data/knowledgebase.xlsx", sheetName = "priceBatteryPower20112012", append = TRUE, overwrite = TRUE)
  
  
  # broken link
  # Langzeitprofil der E-Bike-Käuferschaft in Basel:
  # LangzeitprofilEBikeBasel <- extract_tables("https://www.newride.ch/documents/forschung/NR_BerichtLangzeitprofil2012_EBike_2012_08_28.pdf")
  
  
  # Bundesamt für Statistik, & Bundesamt für Raumplanung. (2012). Mobilität in der Schweiz, Ergebnisse des Mikrozensus Mobilität und Verkehr 2010.
  download.file("https://www.bfs.admin.ch/bfsstatic/dam/assets/291639/master", destfile = "../../data/MikrozensusMobilitaet.xlsx")
  
  
  
  download.file("https://www.bfs.admin.ch/bfsstatic/dam/assets/6046027/master", destfile = "../../data/AltersaufbauCH.xlsx")
  altersverteilung <- data.frame()
  for(sheetName in  2013:2017){
    altersverteilungneu <- read.xlsx(xlsxFile = "../../data/AltersaufbauCH.xlsx", sheet = as.character(sheetName), rows = 9:94, cols = c(1,2,3), colNames = FALSE)
    names(altersverteilungneu) <- c("Alter", "Maenner", "Frauen")
    altersverteilungneu$Jahr = sheetName
    altersverteilung <- rbind(altersverteilung, altersverteilungneu)
  }
  altersverteilung$Maenner <- as.numeric(gsub("(|)", "", altersverteilung$Maenner))
  altersverteilung$Frauen <- as.numeric(gsub("(|)", "",  altersverteilung$Frauen))
  altersverteilung <- altersverteilung %>% 
    gather(Gender, N, c(-Jahr, -Alter)) %>% 
    mutate(N = round(N * 1000,0)) %>% 
    group_by(Jahr, Alter) %>% 
    mutate(Total = sum(N, na.rm = TRUE)) %>% 
    spread(Gender, N) %>% 
    arrange(Jahr, Alter) %>% 
    group_by(Jahr) %>% 
    mutate(cumPct = cumsum(Total)/sum(Total)) %>% 
    select(Jahr, Alter, Maenner, Frauen, Total, cumPct)
  write.excel(data = altersverteilung, file = "../../data/knowledgebase.xlsx", sheetName = "altersverteilung20132017", append = TRUE, overwrite = TRUE)
}

factorsPrimEnergyGHGPotential <- read.excel(file = "../../data/knowledgebase.xlsx", sheetName = "factorsPrimEnergyGHGPotential")

effectEBikeVSregularBike <- read.excel(file = "../../data/knowledgebase.xlsx", sheetName = "effectEBikeVSregularBike")

CHMarket <- read.excel(file = "../../data/knowledgebase.xlsx", sheetName = "velosuisseGesamtmarkt") %>% 
  filter(Segment != "" & Total != "") %>% 
  mutate(Total = gsub("'", "", Total, fixed=TRUE),
         Diff = gsub("'", "", Diff, fixed=TRUE),
         Total2016 = as.numeric(str_split_fixed(Total, " ", 2)[,1]),
         Total2017 = as.numeric(str_split_fixed(Total, " ", 2)[,2]),
         DiffAbsolut = as.numeric(str_split_fixed(Diff, " ", 2)[,1])) 

ecfCapitalCities <- read.excel(file = "../../data/knowledgebase.xlsx", sheetName = "ecfCapitalCities")

ecfMarketEU <- read.excel(file = "../../data/knowledgebase.xlsx", sheetName = "ecfMarketEU")


mikrozensusCHTagesdistanz <- read.excel(file =  "../../exercise1/knowledge_base/MikrozensusMobilitaet.xlsx", sheetName = "Tagesdistanz")[c(-1,-2,-3),c(2,3,4)] %>% 
    fill(X2) %>%
  mutate(mode = X2,
         goal = X3,
         km = as.numeric(X4)) %>% 
  select(mode, goal, km)

priceBatteryPower20112012 <-  read.excel(file = "../../data/knowledgebase.xlsx", sheetName = "priceBatteryPower20112012") %>% 
  mutate(Battery.powerkWh = ifelse(Battery.powerkWh > 100, Battery.powerkWh/1000, Battery.powerkWh))

pricesConventionalBike <- read.excel(file = "../../data/knowledgebase.xlsx", sheetName = "pricesConventionalBike")

altersverteilung <- read.excel(file = "../../data/knowledgebase.xlsx", sheetName = "altersverteilung20132017")
```


\includegraphics[page=126,width=\paperwidth]{../../exercise1/knowledge_base/Verbreitung_Auswirkung_EBikes_CH.pdf}

Source: Verbreitung und Auswirkungen von E-Bikes in der Schweiz: http://www.news.admin.ch/NSBSubscriber/message/attachments/36764.pdf


```{r printKnowledge_CHMarket, echo=FALSE}
kable(CHMarket, caption = "CH Market 2016 - 2017")
```

Source: Annual statistics velosuisse: http://www.velosuisse.ch/de/statistik_aktuell.html



```{r printKnowledge_factorsPrimEnergyGHGPotential, echo=FALSE}
kable(factorsPrimEnergyGHGPotential, caption = "Factors Primary Energy and GHG Potential")
```

Source: Verbreitung und Auswirkungen von E-Bikes in der Schweiz: http://www.news.admin.ch/NSBSubscriber/message/attachments/36764.pdf, Figure 2-15 page 49


\newpage
```{r printKnowledge_effectEBikeVSregularBike, echo=FALSE, paged.print=TRUE}
effectEBikeVSregularBike_tab <- effectEBikeVSregularBike %>% 
  select(-FahrradTouren)
names(effectEBikeVSregularBike_tab) <- gsub("Arbeitsweg", "", names(effectEBikeVSregularBike_tab), fixed=TRUE)

kable(effectEBikeVSregularBike_tab, caption = "Effect of EBike for commuting on regular bike (% Answers)")

```

Source: Verbreitung und Auswirkungen von E-Bikes in der Schweiz: http://www.news.admin.ch/NSBSubscriber/message/attachments/36764.pdf, Figure 3-45, 3-46 page 105, 106


\newpage
```{r printKnowledge_ecfCapitalCities, echo=FALSE}
kable(ecfCapitalCities, caption = "European Cycling Federation EU capitals")
```

Source: Webpage European Cyclist Federation (ECF): https://ecf.com/resources/cycling-facts-and-figures/capital-cities



```{r printKnowledge_ecfMarketEU, echo=FALSE}
kable(ecfMarketEU, caption = "European Cycling Federation EU market")
```

Source: European bicycle market analysis 2015: https://ecf.com/sites/ecf.com/files/CONEBI%20market%20report%20analysis%202016_1.pdf, Table 1 page 17



```{r printKnowledge_pricesConventionalBike, echo=FALSE}
kable(pricesConventionalBike, caption = "Prices conventional bikes")
```

Source: Electrification of road transport – an analysis of the economic performance of electric two- wheelers: https://dspace.library.uu.nl/bitstream/handle/1874/275936/Thesis%20P.W.K.%20Dekker%2012%20May%202013.pdf?sequence=1, Table 10 page 50


\newpage
```{r printKnowledge_priceBatteryPower20112012, echo=FALSE}
kable(priceBatteryPower20112012, caption = "Prices and Battery Power 2011 - 2012")
```

Source: Electrification of road transport – an analysis of the economic performance of electric two- wheelers: https://dspace.library.uu.nl/bitstream/handle/1874/275936/Thesis%20P.W.K.%20Dekker%2012%20May%202013.pdf?sequence=1, Table 12, 13 page 50 -52


\newpage	
```{r printKnowledge_mikrozensusCHTagesdistanzSummary, echo=FALSE}

mikrozensusCHTagesdistanzSummary <- mikrozensusCHTagesdistanz %>% 
  filter(!grepl("Anzahl", mode, fixed=TRUE)) %>% 
  group_by(mode) %>%
  summarise(MeanDistKm = round(mean(km, na.rm=TRUE),2),
            TotalDistKM = round(sum(km, na.rm=TRUE),1))

kable(mikrozensusCHTagesdistanzSummary, caption = "Mikrozensus CH: Tagesdistanz")
```

Source: Bundesamt für Statistik, & Bundesamt für Raumplanung. (2012). Mobilität in der Schweiz, Ergebnisse des Mikrozensus Mobilität und Verkehr 2010. (http://www.portal-stat.admin.ch/mz10/files/de/00.xml), Sheet ''Tagesdistanz''


\newpage	
```{r printKnowledge_altersverteilungSummary, echo=FALSE}

altersverteilungSummary <- altersverteilung %>%
  group_by(Jahr) %>% 
  summarise(unter50 = sum(ifelse(Alter <= 50, Total, 0)),
         ueber50 = sum(ifelse(Alter > 50, Total, 0))) %>%
  mutate(Total = unter50 + ueber50,
         pctUeber50 = round(ueber50/Total * 100, 0))

kable(altersverteilungSummary, caption = "Development proportion of population over 50 years old")

```



Source: Bundesamt für Statistik, & Bundesamt für Raumplanung. (2012). Altersaufbau der ständigen Wohnbevölkerung ab 15 Jahren, nach Migrationsstatus und Geschlecht, 2013-2017 . (https://www.bfs.admin.ch/bfs/de/home/statistiken/bevoelkerung/alterung.assetdetail.6046027.html)



Other Source:
Schwegler, R., Iten, R., Spescha, G., & Schäppi, B. (2015). Umfrage Grüne Wirtschaft und Klima. Technischer Bericht zur Konzeptionierung. Bundesamt für Umwelt BAFU.

	
### Actors to be involved
	-
	
### Target group
	-

\newpage
## 2.1 Impact factors
### preliminary identification:
```{r prelimIF, echo=FALSE}
# load csv and write its content to excel
if(FALSE & file.exists("../../data/prelimImpactFactors.csv")){
  prelimImpactFactors <- read.csv("../../data/prelimImpactFactors.csv", stringsAsFactors = FALSE)
  prelimImpactFactors$Indicator <- NA
  prelimImpactFactors$Current_Stat <- NA
  
  dataXlsxName <- "../../data/exerciseData.xlsx"
  if(!file.exists(dataXlsxName)){
    write.excel(data = prelimImpactFactors, file = "../../data/exerciseData.xlsx", sheetName = "preliminaryImpactFactors", rNames = FALSE)
  } 
  if(file.exists(dataXlsxName) & "preliminaryImpactFactors" %in% excel.sheets(dataXlsxName)){
    write.excel(data = prelimImpactFactors, file = "../../data/exerciseData.xlsx", sheetName = "preliminaryImpactFactors", rNames = FALSE, append = TRUE, overwrite = TRUE)
  }
  if(file.exists(dataXlsxName) & !("preliminaryImpactFactors" %in% excel.sheets(dataXlsxName))){
    write.excel(data = prelimImpactFactors, file = "../../data/exerciseData.xlsx", sheetName = "preliminaryImpactFactors", rNames = FALSE, append = TRUE)
  }
}
# and use the excel
prelimImpactFactors <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "preliminaryImpactFactors")

prelimImpactFactorsList <- as.character(prelimImpactFactors$ImpactFactor)
cols <- 3
nrowPerCol <- ceiling(length(prelimImpactFactorsList)/cols)
prelimImpFac <- c(prelimImpactFactorsList, rep("", each = cols * nrowPerCol - length(prelimImpactFactorsList)))
prelimImpFacTab <- cbind(prelimImpFac[1:nrowPerCol], prelimImpFac[(nrowPerCol+1):(2*nrowPerCol)], prelimImpFac[(2*nrowPerCol+1):(3*nrowPerCol)])
kable(prelimImpFacTab, caption = "Names of preliminary impact factors")


```
		

	Impact factors structuring, clustering, relevance assessment
	
	-
	-
	
	Impact factors selection
	
```{r finalIF, echo=FALSE}
finalImpactFactors <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "finalImpactFactors")
finalImpactFactorsList <- as.character(finalImpactFactors$ImpactFactor)
cols <- 3
nrowPerCol <- ceiling(length(finalImpactFactorsList)/cols)
impFac <- c(finalImpactFactorsList, rep(NA, each = cols * nrowPerCol - length(finalImpactFactorsList)))
impFacTab <- cbind(impFac[1:nrowPerCol], impFac[(nrowPerCol+1):(2*nrowPerCol)], impFac[(2*nrowPerCol+1):(3*nrowPerCol)])
# kable(impFacTab, caption = "Names of seleceted impact factors")
kable(finalImpactFactors[, c("ImpactFactor", "Indicator", "Current_Stat")], caption="Description of seleceted impact factors")
```

\newpage	
## 2.2 Impact assessment
	
```{r updateImpactMatrix, echo=FALSE, warning=FALSE}

mkeIFM <- function(IFList){
  iMatrix <- matrix(0, nrow = length(IFList), ncol=length(IFList))
  rownames(iMatrix) <- IFList
  colnames(iMatrix) <- IFList
  impactMatrix <- as.data.frame(iMatrix)
  for(i in 1:nrow(impactMatrix)){
    impactMatrix[i,i] <- NA
  }
}


finalImpactFactorsList <- as.character(finalImpactFactors$ImpactFactor)

dataXlsxName <- "../../data/exerciseData.xlsx"
if(file.exists(dataXlsxName) & !("impactMatrix" %in% excel.sheets(dataXlsxName))){
  impactMatrix <- mkeIFM(IFList = finalImpactFactorsList)
  write.excel(data = impactMatrix, file = "../../data/exerciseData.xlsx", sheetName = "impactMatrix", append = TRUE, rNames = TRUE)
}

# # write updatet matrix if there are new impactFactors
# if(file.exists("../../data/exerciseData.xlsx") & ("impactMatrix" %in% excel.sheets(dataXlsxName))){
#   oldImpactMatrix <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "impactMatrix")
#   rownames(oldImpactMatrix) <- oldImpactMatrix[,1]
#   oldImpactMatrix <- oldImpactMatrix[,-1]
#   existingIFIndex <- which(rownames(oldImpactMatrix) %in% finalImpactFactorsList)
#   newIFNames <- finalImpactFactorsList[!(finalImpactFactorsList %in% rownames(oldImpactMatrix))]
#   newImpactMatrix <- oldImpactMatrix[existingIFIndex, existingIFIndex]
#   if(length(newImpactMatrix) == 0){
#     newImpactMatrix <- mkeIFM(finalImpactFactorsList)
#   } else {
#     for(newIF in newIFNames) {
#       newData <- 0
#       newImpactMatrix <- cbind(newImpactMatrix, newData)
#       colnames(newImpactMatrix)[ncol(newImpactMatrix)] <- newIF
#       newImpactMatrix <- rbind(newImpactMatrix, newData)
#       rownames(newImpactMatrix)[nrow(newImpactMatrix)] <- newIF
#       newImpactMatrix[nrow(newImpactMatrix), ncol(newImpactMatrix)] <- NA
#     }
#   }
#   write.excel(data = impactMatrix, file = "../../data/exerciseData.xlsx", sheetName = "impactMatrix", append = TRUE, overwrite = TRUE, rNames = TRUE)
# 
# }
```

```{r impactMatrix, echo=FALSE, warning=FALSE}
# always load from excel
impactMatrix <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "impactMatrix")
names(impactMatrix)[1] <- "ImpactFactor"
impactMatrixPlot <- impactMatrix %>% 
  gather(Dependend, value, -ImpactFactor) %>% 
  mutate(value = factor(value)) %>% 
  mutate(value = factor(value),
         ImpactFactor = factor(ImpactFactor),
         ImpactFactor = factor(ImpactFactor, levels = rev(levels(ImpactFactor)))) %>%
  filter(!is.na(value)) %>% 
  ggplot(aes(Dependend, ImpactFactor, fill=value)) +
  geom_tile() +
  scale_fill_manual(values = c("0" = mc(5), "1" = mc(4),"2" = mc(1)), na.value=mc(7))+
  geom_text(aes(label = value)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

impactMatrixPlot
  
```
	
## 2.3 Impact analysis

```{r activityPassivityScores, echo=FALSE}
impactM <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "impactMatrix")
rownames(impactM) <- impactM[,1]
impactM <- impactM[,-1]
impactM$ActivityScore <- rowSums(impactM, na.rm = TRUE)
PassivityScore <- colSums(impactM, na.rm = TRUE)
impactM <- rbind.data.frame(impactM, PassivityScore= PassivityScore)
names(impactM) <- gsub(".", " ", names(impactM), fixed=TRUE)

shortNames <- lapply(str_extract_all(names(impactM), boundary("word")), FUN = substr, start = 1, stop =1)
sNames <- character(0)
for(i in 1: length(shortNames)){
  sNames <- c(sNames, paste0(shortNames[[i]], collapse = "."))
}

names(impactM) <- sNames

# kable(impactM, caption = "Impact Factor matrix")
if(!require("kableExtra")){
  install.packages("kableExtra")
  require("kableExtra")
}

kable(impactM, caption = "Impact Factor matrix", "latex", booktabs = T, escape = F, col.names = linebreak(gsub(" ", "\n",names(impactM)))) %>%
  # kable_styling(font_size = 8) %>%
landscape()

```

### System grid
```{r plotSystemGrid, echo=FALSE}
impactM <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "impactMatrix")
rownames(impactM) <- impactM[,1]
impactM <- impactM[,-1]
impactM$ActivityScore <- rowSums(impactM, na.rm = TRUE)
PassivityScore <- colSums(impactM, na.rm = TRUE)
impactM <- rbind.data.frame(impactM, PassivityScore= PassivityScore)

sysGridData <- impactM
sysGridData$PassivityScore <- as.numeric(t(sysGridData[nrow(sysGridData),]))
sysGridData$ImpactFactor <- rownames(sysGridData)
sysGridData <- sysGridData %>%
  select(ImpactFactor, ActivityScore, PassivityScore) %>%
  filter(!(ImpactFactor %in% c("PassivityScore")))
  


sysGridData %>% 
  ggplot(aes(PassivityScore, ActivityScore, label=ImpactFactor)) +
  geom_point() +
  geom_hline(aes(yintercept = mean(PassivityScore))) +
  geom_vline(aes(xintercept = mean(ActivityScore))) +
  geom_text_repel()


```

### Feedback loops


###System structure
	
\newpage	
## 3.1 Future level definition

Goal: 3 future levels (ni = 1,2,3) for each impact factor d_i

 * 1 future level outlining trend extrapolation
 * Two future levels outlining extreme developments (min. vs. max.)
 
```{r levelDefinitions, echo=FALSE}
impactFactorLevels <- finalImpactFactors %>% 
  mutate(trend = NA,
         min = NA,
         max = NA
         )
if(file.exists("../../data/exerciseData.xlsx") & ("impactFactorLevels" %in% excel.sheets(dataXlsxName))){
  oldImpactFactorLevels <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "impactFactorLevels")
  newImpactFactorLevels <- impactFactorLevels[!(impactFactorLevels$ImpactFactor %in% oldImpactFactorLevels$ImpactFactor),]
  remainingImpactFactorLevels <- oldImpactFactorLevels[oldImpactFactorLevels$ImpactFactor %in% impactFactorLevels$ImpactFactor,]
  
  impactFactorLevels <- rbind(remainingImpactFactorLevels, newImpactFactorLevels)
  write.excel(impactFactorLevels, file = "../../data/exerciseData.xlsx", sheetName = "impactFactorLevels", append = TRUE, overwrite = TRUE, rNames = FALSE )
} else {
  write.excel(impactFactorLevels, file = "../../data/exerciseData.xlsx", sheetName = "impactFactorLevels", append = TRUE, rNames = FALSE )
}

impactFactorLevels <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "impactFactorLevels")


kable(impactFactorLevels[,c("ImpactFactor", "Indicator", "Current_Stat", "trend", "min","max")], caption = "Description of impact factor levels")
```

\newpage


```{r extrapolPrices, echo=FALSE}

expolPrices <- priceBatteryPower20112012 %>% 
  group_by(Year) %>% 
  summarise(N = n(),
            amPrice = median(PriceEur, na.rm = TRUE),
            sdPrice = sd(PriceEur, na.rm = TRUE),
            sePrice = sdPrice/N^0.5,
            amPower = mean(Battery.powerkWh, na.rm = TRUE),
            sdPower = sd(Battery.powerkWh, na.rm = TRUE),
            sePower = sdPower/N^0.5,
            amPricePerPower = mean(Battery.powerkWh/PriceEur, na.rm = TRUE),
            sdPricePerPower = sd(Battery.powerkWh/PriceEur, na.rm = TRUE),
            sePricePerPower = sdPricePerPower/N^0.5) %>% 
  select(Year, amPrice, amPricePerPower, sdPricePerPower)

lmPricePower <- lm(data =priceBatteryPower20112012, PriceEur ~ Year)

kable(expolPrices, caption="Development price per kWh")
```

From:
Electrification of road transport – an analysis of the economic performance of electric twowheelers:

"We find that e-bike prices will decrease by 3% from €1614,- ± 674 in 2012 to €1561,- ± 609,- in 2025"


```{r extrapoPrices2, echo = FALSE}
CHFproEUR <- 1.14

dekker <- data.frame(year = c(2012, 2025),
                     CHF = c(1614 * CHFproEUR, 1561 * CHFproEUR),
                     sdCHF = c(674 * CHFproEUR, 609 * CHFproEUR))

dekkerTrend <- (dekker[2,2] - dekker[1,2])/13
dekkerWorst <- 0
dekkerBest <- (dekker[2,2]  - (dekker[1,2] + 0.5*dekker[1,3]))/13

simEBikePrices <- data.frame(year = c(2012:2050)) %>% 
  mutate(trendCHF = dekker[1,2] + dekkerTrend * (year - 2012),
         worstCHF = dekker[1,2] + dekkerWorst * (year - 2012),
         bestCHF = dekker[1,2] + dekkerBest * (year - 2012)) 


```

```{r extrapoPrices2Table, echo=FALSE}

simEBikePricesTable <- simEBikePrices %>% 
  filter(year %%  10 == 0)

kable(simEBikePricesTable, caption = "Simulation linear trends price EBike")
```


```{r extrapoPrices2Plot, echo=FALSE, fig.cap="Simulation linear trends price EBike"}

simEBikePrices %>% 
  gather(level, value, -year) %>% 
  ggplot(aes(year, value, color=level)) +
  geom_path() +
  expand_limits(y = 0) +
  labs(y = "E Bike prices [CHF]")
```

\newpage


```{r altersverteilungModel, echo=FALSE}

lmMod <- lm(pctUeber50 ~ log(Jahr-2012), data=altersverteilungSummary)
iCept <- lmMod$coefficients[1]
slope <- lmMod$coefficients[2]
se <- summary(lmMod)$coefficients[2,2]

altersverteilungModelPredictions = data.frame(Jahr = 2013:2050) %>% 
  mutate(AnteilUeber50_trend = iCept +  slope * log(Jahr-2012),
         AnteilUeber50_min = iCept + (slope- 1.96*se) * log(Jahr-2012),
         AnteilUeber50_max = iCept + (slope + 1.96*se) * log(Jahr-2012))

```

```{r printKnowledge_altersverteilungModelSummary, echo=FALSE}

altersverteilungModelPredictionsSummary <- round(altersverteilungModelPredictions,0) %>%
  filter(Jahr %% 10 == 0) 

kable(altersverteilungModelPredictionsSummary, caption = "Prediction age distribution proportion over 50 [\\%]")

```

```{r printKnowledge_altersverteilungPredictionPlot, echo=FALSE, fig.cap="Prediction age distribution proportion over 50"}
ggplot(altersverteilungModelPredictions) +
  geom_path(aes(Jahr, AnteilUeber50_trend)) +
  geom_path(aes(Jahr, AnteilUeber50_min)) +
  geom_path(aes(Jahr, AnteilUeber50_max)) +
  geom_point(aes(x=Jahr, y = pctUeber50), data=altersverteilungSummary, color="blue") +
  labs(y="Anteil ueber 50 Jaehriger [%]")


```

## 3.2 Consistency assessment

```{r consostencyAssessment, echo=FALSE}

impactFactorLevelsLong <- impactFactorLevels %>% 
  select(ImpactFactor, trend, min, max) %>%
  gather(level, value, -ImpactFactor) %>% 
  arrange(ImpactFactor)

ifColsNames <- paste(impactFactorLevelsLong$ImpactFactor, impactFactorLevelsLong$value, sep="_")
  ifCols <- data.frame(matrix(rep(rep(NA, each=length(ifColsNames)), each=length(ifColsNames)), ncol=length(ifColsNames)))
  names(ifCols) <- ifColsNames
  
impactFactorConsistency <- impactFactorLevelsLong[,c("ImpactFactor", "value")] %>% 
  bind_cols(ifCols)

if(file.exists("../../data/exerciseData.xlsx") & ("impactFactorConsistency" %in% excel.sheets(dataXlsxName))){
  oldImpactFactorConsistency <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "impactFactorConsistency")



  for(i in 1:nrow(impactFactorConsistency)){
    for(j in 1:ncol(impactFactorConsistency)){
      impactFactorConsistency[1:i,j] <- NA
      impF <-  impactFactorConsistency[i,]$ImpactFactor
      val <- impactFactorConsistency[i,]$value
      iOldCindex <- which(oldImpactFactorConsistency$ImpactFactor == impF & oldImpactFactorConsistency$value == val)
      jOldCindex <- which(names(oldImpactFactorConsistency) == names(impactFactorConsistency[j]))
      if(length(iOldCindex) > 0 & length(jOldCindex) > 0){
        impactFactorConsistency[i,j] <- oldImpactFactorConsistency[iOldCindex, jOldCindex]
      }
    }
  }
  write.excel(impactFactorConsistency, file = "../../data/exerciseData.xlsx", sheetName = "impactFactorConsistency", append = TRUE, overwrite = TRUE, rNames = FALSE )

} else {
write.excel(impactFactorConsistency, file = "../../data/exerciseData.xlsx", sheetName = "impactFactorConsistency", append = TRUE, rNames = FALSE )
}

impactFactorConsistency <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "impactFactorConsistency")
```

## 3.3 Scenario construction

```{r levelSelection, echo=FALSE}

impactFactorConsistent <- impactFactorConsistency %>% 
  gather(ImpactFactor2, level, c(-ImpactFactor,-value)) %>%
  filter(!is.na(level) & level >= 0) %>%
  mutate(ImpactFactor2Name = str_split_fixed(ImpactFactor2, "_",2)[,1],
         ImpactFactor2Level = str_split_fixed(ImpactFactor2, "_",2)[,2]) %>% 
  select(ImpactFactor1 = ImpactFactor, ImpactFactor1Level = value, ImpactFactor2 = ImpactFactor2Name, ImpactFactor2Level, score = level)


mostConsistent <- impactFactorConsistent %>%
  arrange(score) %>% 
  group_by(ImpactFactor1, ImpactFactor2) %>% 
  summarise(level1 = first(ImpactFactor1Level),
            level2 = first(ImpactFactor2Level))



```

## 4.1 Scenario selection

```{r scenarioSelecetion, echo=FALSE}

selectedScenarios <- read.excel(file = "../../data/exerciseData.xlsx", sheetName = "selectedScenarios")

kable(selectedScenarios, caption="Impact Factor levels for the seleceted scenarios")
```


## 4.2 Scenario description and interpretation

### Requirements

```{r scenrioRequirements, echo=FALSE}
scenarioInformation <- data.frame(
  Parameter = c("Market penetration","E-bike transport","Battery technology","Other transport"),
  Description = c("Market penetration", "Kilometers driven per e-bike annually","Share of different battery technologies used in e-bikes sold annually", "Kilometers driven by other means of transport (annually)"),
  PhysicalUnit = c("[no. per year]", "[km per year]", "[% per year]", "[km per year]"))

kable(scenarioInformation, caption = "Information requirements for scenarios")
```


  * Impact factors and their dynamic relevance
  * Final scenarios and their rationale (storyline)
  * Parameters values indicated above for each scenario as a time series until 2050 (cf. Table above)
